---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NMproject

<!-- badges: start -->
[![Travis-CI Build Status](https://travis-ci.org/tsahota/NMproject.svg?branch=master)](https://travis-ci.org/tsahota/NMproject)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tsahota/NMproject?branch=master&svg=true)](https://ci.appveyor.com/project/tsahota/NMproject)
[![Coverage Status](https://coveralls.io/repos/github/tsahota/NMproject/badge.svg?branch=master)](https://coveralls.io/github/tsahota/NMproject?branch=master)
[![Codecov test coverage](https://codecov.io/gh/tsahota/NMproject/branch/master/graph/badge.svg)](https://codecov.io/gh/tsahota/NMproject?branch=master)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![CRAN status](https://www.r-pkg.org/badges/version/NMproject)](https://CRAN.R-project.org/package=NMproject)
<!-- badges: end -->

Script based 'NONMEM' model development in RStudio.

- NONMEM code library
- End-to-end script based model development workflows
- Scale to groups of runs and complex workflows
- 100% flexibility through tracked manual edits to model files
- Customisable to multiple infrastructure types

[Two-minute YouTube summary](https://www.youtube.com/watch?v=b7oBb6QZub8)
*WARNING: this is the prototype "alpha" interface. The new interface has a different syntax and is not backwards compatible. Video is to be replaced*

<!-- old pictures
<img src=https://user-images.githubusercontent.com/18026277/26879195-79b6f4c0-4b90-11e7-8228-01b117e64a12.png width=24.6% /><img src=https://user-images.githubusercontent.com/18026277/26879231-a046cfc0-4b90-11e7-9dbf-666086f32b9d.png width=24.5% /><img src=https://user-images.githubusercontent.com/18026277/26879238-a4a94fc0-4b90-11e7-8e8f-1b12a03f912d.png width=24.5% /><img src=https://user-images.githubusercontent.com/18026277/26879240-a7a53ebe-4b90-11e7-80fa-74bef643db29.png width=24.5% />
-->

## Installation

<!-- CRAN install instructions

You can install the released version of NMproject from [CRAN](https://CRAN.R-project.org) with:

```{r eval = FALSE}
install.packages("NMproject")
```

-->

To install the latest version of NMproject from [GitHub](https://github.com/):

```{r eval = FALSE}
if(!require("devtools")) install.packages("devtools")
devtools::install_github("tsahota/NMproject")
```

To install a specific release (e.g. v0.5.1) on [GitHub](https://github.com/) 
use the following command:

```{r eval = FALSE}
devtools::install_github("tsahota/NMproject@v0.5.1")
```

Load the package with 

```{r eval = FALSE}
library(NMproject)
```

## Getting started with NMproject

Two options:

1. Running the
[demo](https://tsahota.github.io/NMproject/articles/NMproject.html#demo-learn-by-doing-1)
is easiest way to familiarise your with NMproject.
2. Reading the 
[vignette](https://tsahota.github.io/NMproject/articles/NMproject.html).

## Code snippets

Use of pipes, `%>%`, make it easy to code sequences of operations to model 
objects.

Following snippet adds covariates to model object, `m2`:

- create a separate child control file
- add a covariate relationship to it (using PsN SCM syntax)
- run

```{r eval=FALSE}

m2WT <- m2 %>% child() %>%
  add_cov(param = "CL", cov = "WT", state = "power") %>%
  run_nm()

```

Graphical RStudio 'Addins' exist for reviewing the changes that functions like
`add_cov()` make before execution and performing `nm_tran()` checks.

For more complex operations use fully tracked manual edits.

Apply fully customisable diagnostic reports to one or multiple objects with
`nm_render()` like so:

```{r eval = FALSE}

c(m1, m2) %>% nm_render("Scripts/basic_gof.Rmd")
## Saves html diagnostic reports in "Results" directory

```

The template `Scripts/basic_gof.Rmd` can also be run as an R notebook for
interactively customising to your specific model evaluation criteria.

Here's a snippet for producing PPCs and VPCs:

- create a new (child) control stream
- updating initial estimates to final estimates
- convert it to a simulation control file
- run
- generate customised PPCs and VPCs from the outputs

```{r eval=FALSE}
m2s <- m2 %>% child(run_id = "m2s") %>%
  update_parameters(m2) %>%
  convert_to_simulation(subpr = 50) %>%
  run_nm()

m2s %>% nm_render("Scripts/basic_vpc.Rmd")
m2s %>% nm_render("Scripts/basic_ppc.Rmd")

```

Advanced functionality enables groups of runs to be handled with the same
concise syntax (no loops):

- create 5 child runs
- Perturb the initial estimates of \$THETA and \$OMEGA
- move them to their own subdirectory
- run them all

```{r eval = FALSE}

m1rep <- m1 %>% child(1:5) %>% 
  init_theta(init = rnorm(init, mean = init, sd = 0.3)) %>%
  init_omega(init = runif(init, min = init/2, max = init/2)) %>%
  run_in("Models/m1_perturb_inits") %>%
  run_nm()

```

See the
[vignette](https://tsahota.github.io/NMproject/articles/NMproject.html) for more
examples
