---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NMproject

<!-- badges: start -->
[![Travis-CI Build Status](https://travis-ci.org/tsahota/NMproject.svg?branch=master)](https://travis-ci.org/tsahota/NMproject)
[![Coverage Status](https://coveralls.io/repos/github/tsahota/NMproject/badge.svg?branch=master)](https://coveralls.io/github/tsahota/NMproject?branch=master)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tsahota/NMproject?branch=master&svg=true)](https://ci.appveyor.com/project/tsahota/NMproject)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![Codecov test coverage](https://codecov.io/gh/tsahota/NMproject/branch/master/graph/badge.svg)](https://codecov.io/gh/tsahota/NMproject?branch=master)
<!-- badges: end -->

*This is the newer beta interface.  To use the older interface, install v0.3.2 (see releases)*

The intent of NMproject is to allow for fully script based NONMEM model development:

* Code management:
  * Code library: NONMEM template scripts to speed creation of NONMEM runs and facilitate adherence to best practices.
  * Standardised, version controlled, directory structure for all NONMEM users.
* Script based model development:
  * Instead of command line interface (CLI) or graphical user interface (GUI) use a script based interface
  * End-to-end R workflows that can be repeated, reused and documented in R markdown
  * Convenience functions for routine NONMEM control file edits
  * Tracked manual edits for everything else to ensure full flexibility
  * Fully customizable r markdown templates for goodness of fits, vpcs, ppcs, bootstraps
  * Vectorized model objects - interact with groups of NONMEM runs as if they were one (no loops needed)
  * Shiny interface for real time run tracking in grid environments

History: NMproject was previously an AstraZeneca project.  It is being reimplemented here as a community version to be compatible with a variety of architectures (standalone NONMEM and a variety of grid submission systems)

Two-minute Youtube summary (alpha interface): https://www.youtube.com/watch?v=b7oBb6QZub8

<!-- old pictures
<img src=https://user-images.githubusercontent.com/18026277/26879195-79b6f4c0-4b90-11e7-8228-01b117e64a12.png width=24.6% /><img src=https://user-images.githubusercontent.com/18026277/26879231-a046cfc0-4b90-11e7-9dbf-666086f32b9d.png width=24.5% /><img src=https://user-images.githubusercontent.com/18026277/26879238-a4a94fc0-4b90-11e7-8e8f-1b12a03f912d.png width=24.5% /><img src=https://user-images.githubusercontent.com/18026277/26879240-a7a53ebe-4b90-11e7-80fa-74bef643db29.png width=24.5% />
-->

## Installation

NONMEM, PsN, and Rstudio are required to be installed prior to these steps.  To install a specific release version (e.g. v0.5.1) from [GitHub](https://github.com/):

```{r eval = FALSE}
if(!require("devtools")) install.packages("devtools")
devtools::install_github("tsahota/NMproject@v0.5.1")
```

To install the latest developmental version:

```{r eval = FALSE}
devtools::install_github("tsahota/NMproject")
```

Load the package with 

```{r eval = FALSE}
library(NMproject)
```

## Instructions

* NMprojects are directories where you can work on pharmacometric analysis.
* First, set up a tidyproject with the `make_project("/path/to/project/dir"")` or in RStudio File -> New Project -> New Directory -> New tidyproject
   * See tutorial at https://github.com/tsahota/tidyproject for more information

### Demo

The easiest way to familiarise your with NMproject is to follow through the demo.

First create a new tidyproject (`FILE` -> `New Project` -> `New Directory` -> `New tidyproject`). You'll see a clean analysis directory with empty subfolders.  Run the following in the new R session:

```{r eval = FALSE}
library(NMproject)
setup_nm_demo()
```

This will population the `Scripts` directory with scripts, and deposit the Theopp dataset into the `SourceData` subdirectory.

The scripts are numbered s01.....Rmd, s02....Rmd etc., and are designed to be read and run in order.  This is the best way to familiarize yourself.  The scripts can also be used as template for your own model development

## Example

If we already have a parent run `m1` using `ADVAN2 TRANS1`, we can create a `child()` run that uses `TRANS2` using `subroutine()`:

```{r eval=FALSE}

m2 <- m1 %>% child() %>%
  subroutine(trans = 2) %>%
  run_nm()

m2 %>% nm_render("Scripts/basic_gof.Rmd")
## run reports will be in "Results" directory of analysis project

```

Create a simulation run using `update_parameters()` and `convert_to_simulation()`.  Apply vpc and ppc diagnostics
  
```{r eval=FALSE}
m2s <- m2 %>% child(run_id = "m2s") %>%
              update_parameters(m2) %>%
              convert_to_simulation(subpr = 50) %>%
              run_nm()

m2s %>% nm_render("Scripts/basic_vpc.Rmd")
m2s %>% nm_render("Scripts/basic_ppc.Rmd")

## run reports will be in "Results" directory of analysis project

```

See the vignette for more examples
