---
title: "Configuration & Setup Options"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(NMproject)
```

## Prerequisites

The following need to be installed and available:

- PsN >= 4.4.8
- NONMEM installed with valid license
- RStudio

## First time set up

### Set up code completion

It is highly recommended to set up code completion though as it eases
the scripting process.

```{r eval = FALSE}
setup_code_completion()
```

This will ask for user consent as it will add entries to the user level snippets
file.

### Verify installation 

The following function checks for availability of PsN via `system_nm()`, this is
a requirement for NMproject.  There are also subsequent checks which not
mandatory, but recommended for full functionality.

```{r eval = FALSE}
check_installation()
```

```{r echo = FALSE}
## dummy line to show test passed message
testthat::test_that("dummy", testthat::expect_true(TRUE))
```

Users working on a standalone machines (e.g. laptops and desktops) should expect
these tests to pass and for NMproject to work out of the box (provided the
pre-requisites are met).

## Custom configuration

More complex environments involving multiple servers and grids may require more
configuration.  The options shown here aim to facilitate that.  You may need the
help of an admin.

### NMTRAN checks

If working in a grid environment, it is highly recommended to perform NMTRAN
checks prior to sending jobs off the grid.  NMTRAN check provides immediate
feedback of control file and dataset errors allowing you to address problems
quickly just like if you were on a laptop/desktop environment.  Without this,
having to wait for a job to reach the grid and fail is a much slower (and more
annoying) process.

NMproject tries to guess the location to the NMTRAN file in the NONMEM
installation directory however if this fails and `check_installation()` cannot
detect it, you will need to specify the NMTRAN command/file with the
`nm_tran_command()` function.  See the following for instructions and examples
at the bottom of the help file:

```{r eval = FALSE}
nm_tran_command("/opt/NONMEM/nm75/tr/NMTRAN.exe")
```

To see more examples see `?nm_tran_command`.

This setting will reset upon R session restart, if you want to make it permanent
see [Making configuration changes permanent](#permanent).

### The `system_nm()` command

Under the hood, all calls to PsN proceed through the `system_nm()` command. This
is a user configurable function.  By default it's the same as `system()` on unix
systems and `shell()` on windows.  However if the NONMEM server is remote and
requires connection (e.g. via SSH), users may want to modify the behaviour of
`system_nm()`.  This is possible through setting the `option()` also named
`system_nm`, e.g. the following allows runs the `system()` command on a remote
machine named `clustername`:

```{r eval = FALSE}
options(system_nm=function(cmd,...) {
        system(paste0("ssh -q clustername \"cd $(pwd); ", cmd, "\""),...)
})
```

This setting will reset upon R session restart, if you want to make it permanent see [Making configuration changes permanent](#permanent).

### Custom code library location

NMproject ships with a code library that can be used without configuration.  If
however you want to have a custom code library, the option `code_library_path`
can be used to set this.  For example:

```{r eval = FALSE}
options(code_library_path = "/projects/myorganisation/code_library")
```

To see more information see `?code_library`.

This setting will reset upon R session restart, if you want to make it permanent
see [Making configuration changes permanent](#permanent).

### Custom directory structures

If your organisation prefers a project directory sub-structure which is
different from the standard this can be configured with the function
`nm_default_dirs()`.  For example if your organisation performs NONMEM modelling
in a directory called "NONMEM" and has a set structure for other
sub-directories, you may have something like:

```{r eval = FALSE}
nm_default_dirs(list(
  models = "NONMEM",
  scripts = "RSCRIPTS",
  results = "OUTPUTS",
  source_data = "DATA/SOURCE",
  derived_data = "DATA/DERIVED"
))
```

To see more examples see `?nm_default_dirs`.

This setting will reset upon R session restart, if you want to make it permanent
see [Making configuration changes permanent](#permanent).

### Custom object properties

If you have a custom PsN command you tend to submit for all NONMEM jobs, use the
`nm_default_fields` function.  For example to set a default parafile and modify
the basic `"execute {ctl_name} -dir={run_dir}"` command to use this parafile,
you would use the command:

```{r eval = FALSE}
nm_default_fields(list(
  parafile = "/opt/NONMEM/nm75/run/mpilinux8.pnm"
  cmd = "execute -parafile={parafile} {ctl_name} -dir={run_dir} -nodes={cores}"
))
```

To see more examples see `?nm_default_fields`.

This setting will reset upon R session restart, if you want to make it permanent
see [Making configuration changes permanent](#permanent).

## Making configuration changes permanent {#permanent}

### Via startup scripts

Whether it be a custom NMTRAN path, `system_nm` option or a different default
directory structure, to make the above changes permanent, they need to be
executed every R session.

R has a few mechanisms for achieving this.  See `?.Rprofile` for details. The
following will open your config file:

```{r eval = FALSE}
usethis::edit_r_profile()
```

Then insert the configuration code into this file and restart R.  For example a
custom NMTRAN path can be set permanently to `"/opt/NONMEM/nm75/tr/NMTRAN.exe"`
by including `nm_tran_command("/opt/NONMEM/nm75/tr/NMTRAN.exe")` into the file.

To set the option site wide you can use the file `$R_HOME/etc/Rprofile.site`
(requires administration privileges).

### Via a custom package

If you already have a package you maintain to deliver functions to users, this
package can be used to configure set up options via the `.onLoad` hook.  For
example, the following will configure custom NMTRAN and code library settings on
package load

```{r eval = FALSE}
.onLoad <- function(libname, pkgname){
  ## custom NMTRAN location  
  NMproject::nm_tran_command("/opt/NONMEM/nm75/tr/NMTRAN.exe")
  ## custom code library location
  options(code_library_path = "/projects/myorganisation/code_library")
  ## message to users in interactive sessions
  if (interactive()) usethis::ui_done("NMproject configured")
  
}
```

The advantage of a package is you can also bundle other functions in there for
users to use.

