---
title: "NONMEM execution"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(knitr)
library(NMproject)
library(dplyr)
extdata_path <- system.file("extdata", package = "NMproject")

testfilesloc <- file.path(extdata_path, "theopp")
zip_file <- file.path(extdata_path, "theopp.zip")

files_to_unzip <- unzip(zip_file, list = TRUE)

unzip(zip_file)

dir(all.files = TRUE)
extracted_folders <- dir("theopp", full.names = TRUE)

for(extracted_folder in extracted_folders){
  file.copy(extracted_folder, ".", recursive = TRUE)
}

file.rename("cache", ".cache")
dir(all.files = TRUE)

copied_folders <- basename(gsub("cache", ".cache", extracted_folders))

unlink("theopp", recursive = TRUE)

orig_dir <- getwd()

```

```{r include=FALSE}

m1 <- dir(".cache", full.names = TRUE) %>%
  subset(grepl("m1$",.)) %>%
  {as_nm_list(readRDS(.)$object)}

m2 <- dir(".cache", full.names = TRUE) %>%
  subset(grepl("m2$",.)) %>%
  {as_nm_list(readRDS(.)$object)}

m2WT <- dir(".cache", full.names = TRUE) %>%
  subset(grepl("m2WT$",.)) %>%
  {as_nm_list(readRDS(.)$object)}

m1orig <- m1

## need to skip overwrite because new user will mean directory gets overwriten
overwrite_behaviour("skip") 

```

## Fast NMTRAN checks

PLACEHOLDER: YouTube video showing "nm_tran" RStudio 'Addin' in action.

In a cluster environment it is especially important to be able to diagnose data and control file issues before sending jobs off to the grid.  Otherwise, you often need to wait minutes for errors that can be spotted immediately.  It's therefore highly recommended to make use of the "nm_tran" RStudio 'Addin'.  This will only run NMTRAN checks to find control file syntax errors and dataset errors.  It will not run NONMEM itself.  To use this 'Addin', highlight the above code and select the "nm_tran" RStudio 'Addin' near the top of the RStudio GUI.

## Executing NONMEM

Executing NONMEM simply requires the `run_nm()` function which will often form the last step of the chain and then run the command.

```{r eval=FALSE}

m1 <- new_nm(run_id = "m1",
             based_on = "staging/Models/ADVAN2.mod",
             data_path = "DerivedData/data.csv") %>%
  cmd("execute {ctl_name} -dir={run_dir}") %>%
  fill_input() %>%
  run_nm()

```

By default the execution will take place asynchronously.

## Shiny run monitor

To view all runs in the workspace and track progress:

```{r eval=FALSE}
shiny_nm()
```

PLACEHOLDER: video showing shiny run monitor

## Clean up runs

Directory sizes can become large with running NONMEM.  One strategy is to use the `clean` option in PsN, however if you want to retain temporary files and delete them post-hoc you can use the `ls_tempfiles()` function to list them for deletion.

```{r eval=FALSE}

m1 %>% ls_tempfiles() ## will list all temporary files associate with run m1

## remove all temporary files associated with m1
m1 %>% ls_tempfiles() %>% unlink()

## Not specifying an argument will list temporary files of all runs
ls_tempfiles()

```

## Running NONMEM with parallel processing

Two fields `parafile` and `cores` when combined with a suitable `cmd` can enable runs to be executed in parallel using PsN.

```{r eval=FALSE}
m1 <- new_nm(run_id = "m1",
             based_on = "staging/Models/ADVAN2.mod",
             data_path = "DerivedData/data.csv") %>%
  cmd("execute {ctl_name} -parafile={parafile} -dir={run_dir} -nodes={cores}") %>%
  parafile("/opt/NONMEM/nm750/run/mpilinux8.pnm") %>%
  cores(4) %>%
  run_nm()

```

```{r include = FALSE}
m1 <- m1 %>%
  cmd("execute {ctl_name} -parafile={parafile} -dir={run_dir} -nodes={cores}") %>%
  parafile("/opt/NONMEM/nm750/run/mpilinux8.pnm") %>%
  cores(4)
```

## Parallel efficiency test

Often you'll want to know the right level of parallelisation to run your model to maximise speed without wasting too many resources.  The following creates multiple runes with different levels of parallelisation. We'll just test it across 1, 3, 10, and 30 cores, but this can be any vector.

```{r}

test_cores <- c(1, 3, 10, 30)  ## test these levels of parallelisation

m1c <- m1 %>% ## start with parent m1 again
  child(run_id = test_cores) %>%  ## use "cores" as run_id (object is now length 4)
  run_in("Models/m1_coretest") %>% ## for tidiness run them all in m1_coretest
  cmd("execute {ctl_name} -parafile={parafile} -dir={run_dir} -nodes={cores}") %>% ## parallelized execute
  parafile("/opt/NONMEM/nm75/run/mpilinux8.pnm") %>%
  cores(test_cores) ## and finally set the cores of each models according to test_cores

m1c %>% cmd()

```

Each run is identical apart from the level of parallelisation

```{r eval = FALSE}

## run them all and wait for them to finish
m1c %>% run_nm() %>% wait_finish()

## extract job statistics and plot cores vs Rtime or Ttime to get plots of run
## time and total time vs number of CPUs

m1c %>%
  job_stats() %>%
  ggplot(aes(x = cores, y = Rtime)) + theme_bw() +
  geom_point() +
  geom_line() +
  expand_limits(y = 0)
  
```

```{r echo = FALSE}

dplyr::tibble(cores = test_cores,
              Rtime = c(20.3, 8.1, 3.9, 3.1)) %>%
  ggplot2::ggplot(ggplot2::aes(x = cores, y = Rtime)) + ggplot2::theme_bw() +
  ggplot2::geom_point() +
  ggplot2::geom_line() + 
  ggplot2::expand_limits(y = 0)
  
```

```{r include=FALSE}
## clean up 
getwd()
unlink(copied_folders, recursive = TRUE)
dir(all.files = TRUE)
```
