---
title: "Post-processing & diagnostics"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(knitr)
library(NMproject)
library(dplyr)
extdata_path <- system.file("extdata", package = "NMproject")

testfilesloc <- file.path(extdata_path, "theopp")
zip_file <- file.path(extdata_path, "theopp.zip")

files_to_unzip <- unzip(zip_file, list = TRUE)

unzip(zip_file)

dir(all.files = TRUE)
extracted_folders <- dir("theopp", full.names = TRUE)

for(extracted_folder in extracted_folders){
  file.copy(extracted_folder, ".", recursive = TRUE)
}

file.rename("cache", ".cache")
dir(all.files = TRUE)

copied_folders <- basename(gsub("cache", ".cache", extracted_folders))

unlink("theopp", recursive = TRUE)

orig_dir <- getwd()

```

```{r include=FALSE}

m1 <- dir(".cache", full.names = TRUE) %>%
  subset(grepl("m1$",.)) %>%
  {as_nm_list(readRDS(.)$object)}

m2 <- dir(".cache", full.names = TRUE) %>%
  subset(grepl("m2$",.)) %>%
  {as_nm_list(readRDS(.)$object)}

m2WT <- dir(".cache", full.names = TRUE) %>%
  subset(grepl("m2WT$",.)) %>%
  {as_nm_list(readRDS(.)$object)}

m1orig <- m1

## need to skip overwrite because new user will mean directory gets overwriten
overwrite_behaviour("skip") 

```

## Diagnostics

Use markdown templates to create a customised set of diagnostics to reuse on multiple models.  In the demo an example is shown in `Scripts/basic_gof.Rmd`, but ideally you'll create your own customised version with everything you need to evaluate your model. To create an R markdown diagnostic template go to `FILE` -> `New File` -> `R markdown` -> `From Template` then select from one of the following:

- `model diagnostic`

- `VPC diagnostic`

- `PPC diagnostic`

- `bootstrap results`

*PLACEHOLDER: short YouTube video of opening, customising, saving and running a template.*

A template will appear in the script window for you to customise.  Instructions are the at the top.  Save the file (e.g. as `Scripts/basic_gof.Rmd`) and run in your log script with:

```{r eval=FALSE}
m1 <- m1 %>% nm_render("Scripts/basic_gof.Rmd")
```

It will create the output in `Results` (or `results_dir(m1)`)

Apply a fully editable goodness of fit R markdown template to both runs `m2` and `m3`:

```{r eval=FALSE}

c(m1, m2) %>% nm_render("Scripts/basic_gof.Rmd")

```

Note: `c(m1, m3)` is a vector object of 2 NONMEM runs.  We'll touch more on this later.

Evaluate and compare runs on the fly with the following commands

```{r}
rr(c(m2, m2WT))
plot_iter(m2, skip = 10) ## skip first 10 interations 
covariance_plot(m2)
```

However these are better placed inside templates to enable rapid and consistent re-use.

## Simulation based diagnostics

Create simulation based diagnostics first by running a simulation using `update_parameters()` and `convert_to_simulation()`.  Then use the `ppc diagnostics` and `vpc diagnostics` R markdown templates to generate customisable VPC and PPC diagnostics reports.
  
```{r eval=FALSE}
m2s <- m2 %>% child(run_id = "m2s") %>%
              update_parameters(m2) %>%
              convert_to_simulation(subpr = 50) %>%
              run_nm()

m2s %>% nm_render("Scripts/basic_vpc.Rmd")
m2s %>% nm_render("Scripts/basic_ppc.Rmd")

```

### VPC diagnostics

The `vpc diagnostics` R markdown templates use the `vpc` package to create VPCs.

### PPC diagnostics

The `ppc diagnostics` R markdown templates use the `NMproject` functions to allow easy creation of VPCs.

The example code contains two functions, `idEXPstat` for computing exposure statistics from raw data, and `EXPstat` for computing summaries (median and CV%) of exposure statistics.  These will be applied to both the raw data and simulated datasets to 

```{r echo = FALSE}

## need the derived dataset in there for the PPC code below to work
## this is needed for input_data(m1) etc., to work
d <- read.csv("SourceData/THEOPP.csv", na = ".") %>%
  group_by(ID) %>%
  mutate(WT = na.omit(WT)) %>%
  write_derived_data("DerivedData/data")

source("R/AUC.R")

```

```{r eval = FALSE}

idEXPstat <- function(d, ...){ ## example statistic function
  ## arg = nonmem dataset data.frame
  ## return data.frame with statistic column
  d %>% group_by(ID, ...) %>% filter(is.na(AMT)) %>%
    summarise(
      AUC = AUC(time = TIME, conc = DV),
      CMAX = max(DV, na.rm = TRUE),
      TMAX = TIME[which.max(DV)]
    ) %>%
    tidyr::gather(key = "exposure", value = "statistic", AUC:TMAX) %>%
    ungroup()
  ## must return a dataset that summarises longitudinal dataset into one per ID, and must contain a "statistic" column
}

EXPstat <- function(d, ...){ ## example statistic function
  ## arg = nonmem dataset data.frame
  ## return data.frame with statistic column
  d %>% idEXPstat(...) %>%  ## reuse idEXPstat for individual stats
    ## summarise over study and any other variables (...)
    group_by(exposure, ...) %>%
    summarise(
      median = median(statistic, na.rm = TRUE),
      cv = 100*sd(statistic, na.rm = TRUE)/mean(statistic, na.rm = TRUE)
    ) %>%
    tidyr::gather(key = "type", value = "statistic", median:cv)
  ## must return a dataset that summarises longitudinal dataset into one per "cohort", and must contain a "statistic" column
}

```


```{r, eval = FALSE}

## here msim is nm vector of simulation control files
ppc_data0 <- msim %>% ppc_data(EXPstat)

ppc_data0 %>%
  filter(type %in% "median") %>%
  mutate(exposure = paste(type, exposure, sep = "_")) %>%
  ppc_histogram_plot(exposure)

ppc_data0 %>%
  filter(type %in% "cv") %>%
  mutate(exposure = paste(type, exposure, sep = "_")) %>%
  ppc_histogram_plot(exposure)

```

We can conclude that the model predicts median exposure and variability of exposure well. The remaining plots show stratified versions of these.

## Bootstraps

The package `rsample` can be used to create bootstrap datasets in your initial data manipulation scripts.  The following is an example bootstrap dataset being prepared with stratification on `SEX` and bodyweight `WTC` categorised in two categories  

```{r eval = FALSE}

d <- d %>%
  mutate(WT_C = cut(WT, breaks = 2, labels = FALSE),
         STRATA = paste(SEX, WT_C, sep = "_"))

d_id <- d %>% distinct(ID, STRATA)

set.seed(123)

## create large set of resamples (to enable simulation to grow without ruining seed)
bootsplits <- rsample::bootstraps(d_id, 100, strata = "STRATA")

dir.create("DerivedData", showWarnings = FALSE)
bootsplits %>% saveRDS("DerivedData/bootsplit_data.csv.RData")

```

In a model development script, the following, performs a 100 sample bootstrap of model `m1`

```{r eval=FALSE}

m1_boot <- m1 %>% make_boot_datasets(samples = 100, overwrite = TRUE)

m1_boot$m %>% run_nm()

## the following bootstrap template will wait for results to complete
m1_boot$m %>% nm_list_render("Scripts/basic_boot.Rmd")

```

Results can be viewed in `Results/basic_boot.m1.html`.

## Running VPCs and Bootstraps through PsN

By changing `cmd` to reflect the proper PsN command, you can get PsN to handle the bootstrap or vpc.

```{r eval=FALSE}

m1psnboot <- m1 %>% child(run_id = "m1psnboot") %>%
  cmd("bootstrap {ctl_name} -samples=50 -dir={run_dir}") %>%
  run_nm()

m1psnvpc <- m1 %>% child(run_id = "m1psnvpc") %>%
  cmd("vpc {ctl_name} -samples=50 -dir={run_dir}") %>%
  run_nm()

```


```{r include=FALSE}
## clean up 
getwd()
## added Derived data in here because it was created in the middle of this script
unlink(c(copied_folders, "DerivedData"), recursive = TRUE)
dir(all.files = TRUE)
```
