---
title: "Analysis Directories"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(NMproject)
```

# Creating Analysis Project 

It is recommended to work in structured.  Create a new analysis project via the RStudio menu items: `FILE` -> `New Project` -> `New Directory` -> `New NMproject`.  The underlying function being used to create this analysis project is `nm_create_analysis_project()`.  See documentation for detailed information including how to modify the structure to suit your preferences.

Follow through the instructions, you'll be asked for a location, a name and whether you want to use `renv` to manage project library directories.  See `renv` documentation for more information.

<iframe width="560" height="315" src="https://www.youtube.com/embed/nAkcEFz0RLg" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

Once created you'll see a clean analysis directory with empty subfolders.

Default the subdirectories for model development (these can be modified - see `nm_create_analysis_project()` documentation):

- SourceData : intended for unmodified source datasets entering the analysis project

- DerivedData : intended for cleaned and processed NONMEM ready datasets

- Scripts : intended for all R scripts

- Models : intended for all NONMEM modelling

- Results : intended as default location for run diagnostics, plots and tables

# Creating new R markdown documents

There is nothing mandating the use of R markdown in NMproject.  You can use scripts.  However R markdown documents produce nice shareable model development logs which provide a readable description of what steps were performed and in what order.

It is advisable to always start from a template.  Templates can be accessed in `File` -> `New File` -> `R markdown...` -> `From Template`.  Two to get started with are `NMproject generic` which we'll use as a generic template for data processing and cleaning and `model development` for our model development notebook.
