---
title: "3. The Demo"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(NMproject)
```

The easiest way to familiarise your with NMproject is to run through the demo scripts line by line.  Here's a short YouTube video on running the demo.

<iframe width="560" height="315" src="https://www.youtube.com/embed/nAkcEFz0RLg" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

The video shows that the first step is to create an analysis directory to play in.  If you don't know how to do this, read the [Analysis directories vignette](https://tsahota.github.io/NMproject/articles/a2_analysis_directories.html).

Open your analysis directory RStudio project and run the following to bring in the files:

```{r eval = FALSE}
library(NMproject)
setup_nm_demo()
```

This will population the `Scripts` directory with scripts, and deposit the standard theophylline dataset into the `SourceData` subdirectory.

Default the subdirectories for model development:

- SourceData : intended for unmodified source datasets entering the analysis project

- DerivedData : intended for cleaned and processed NONMEM ready datasets

- Scripts : intended for all R scripts

- Models : intended for all NONMEM modelling

- Results : intended as default location for run diagnostics, plots and tables

After running the above commands, the Scripts directory should contain scripts numbered `s01.....Rmd`, `s02....Rmd` etc.  These are designed to be read, understood and run in order.

You can run these scripts either as notebooks (`Run` -> `Run All`) or as knitted documents via the `Knit` button.  The advantage of notebooks is that code can be re-execute in chunks interactively.  Knitr produces one document but the session is run non-interactively.  For this reason it is best to start off working with interactive notebooks and then once you're happy with the script, to knit it non-interactively.

To create your own markdown documents from scratch, see [here](https://tsahota.github.io/NMproject/articles/a2_analysis_directories.html#creating-new-r-markdown-documents).

