---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
if (file.exists(".Rprofile")) source(".Rprofile")
```
# {{{ Package }}}

<!-- badges: start -->
<!-- badges: end -->

## Objective

The goal of {{{ Package }}} is to ...

## Directory structure

- **{{{scripts}}}**: Location of Script files for data cleaning, analysis and post-processing.
- **{{{models}}}**: Location of model files (e.g. NONMEM, Stan, ...).
- **R**: Optional location of any custom R functions used in scripts.  These can be loaded with `devtools::load_all()` or by sourcing individually via `source("R/functionname.R")`.
- **{{{results}}}**: Location of Results files
- **{{{source_data}}}**: Location of unmodified source datasets
- **{{{derived_data}}}**: Location of analysis-ready datasets.  The production of these should be scripted.

## Automated tests

Any automated tests will be included here.

```{r echo = FALSE}

if (file.exists(file.path("tests", "testthat.R"))){
  devtools::test()
} 

```

## Data sources

The data was obtained from ...

<!--

Include links to databases and contact points to ensure this step can be traced.

-->

## Instructions to reproduce analysis

<!-- 

It is recommended to use a package management system (like renv) for all non-exploratory analysis that may need to be reproduced at future dates

-->

## More information

