---
title: "model development"
author: "`r Sys.info()['user']`"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


```{r setup, include=F}
## DO NOT MODIFY THIS BLOCK (unless you know what you're doing)
library(knitr)
library(rprojroot)
opts_knit$set(root.dir = find_root(has_file(".Rprofile") | 
                                     is_rstudio_project | 
                                     is_r_package | 
                                     is_git_root))
opts_chunk$set(echo = TRUE)
opts_chunk$set(message = TRUE)
```

```{r echo=FALSE,message=FALSE}
## LOAD PACKAGES HERE

library(NMproject)
# library(future)
# future::plan("future::multiprocess", workers = 2)
library(dplyr)

# devtools::load_all()

```

## model development

Demo has already copied a control file "ADVAN2.mod" from code library into staging area.

First parent object takes the most effort to get right. Subsequent child runs will inherit characteristics of parents

```{r parent_ob}

m1 <- new_nm(run_id = "m1",
             based_on = "staging/Models/ADVAN2.mod",
             data_path = "DerivedData/data.csv") %>%
  # See all fields by highlighting the command so far (not including the pipe
  # %>% ) and send it to the console. You can then display m1 so far and see
  # that the field "ctl_name" is "Models/runm1.mod".  Note that the control file
  # will not be written to disk until nonmem is ready to run (the run_nm() step
  # at the end).
  
  # cmd sets the psn command that will be used to run NONMEM (i.e. when we run
  # run_nm()).  Note that the command uses {ctl_name} and {run_dir} which are
  # placeholders for the ctl_name and run_dir fields of the object (can be
  # viewed by print the object so far and by using ctl_name(m1) and run_dir(m1)
  # respectively.
  cmd("execute {ctl_name} -dir={run_dir}") %>%
  # view the control file using show_ctl(m1) and note that $DATA has already
  # been set to the right location, and that $INPUT, $THETA and $SIGMA need
  # filling in.  We could manually edit this (manual_edit addin), but instead
  # we'll do this automatically using three functions fill_input(), init_theta()
  # and init_sigma()
  fill_input() %>%
  init_theta(init = c(0.5, -2.5, -0.5)) %>%
  init_sigma(init = c(0.1, 0.1)) %>%
  # to check fill_input() and init_*() worked as intended, highlight this entire
  # code segment and select addins -> nm_diff.  This is sensible practice after
  # all automatic edits and will display the unix file diff.  Get used to
  # reading these for a concise description of file changes. Then run nm_tran()
  # in a similar way via the addins -> nm_tran().
  run_nm()  # last step is usually to run nonmem.  only at this point will the control file be written to disk.

```

the demo contains a simple post processing template. Generate this from
scratch by clicking New File -> R markdown -> From template -> basic gof. the
default template is very minimal but can be customised and reused by following
instructions

```{r basic_gof}

m1 %>% nm_render("Scripts/basic_gof.Rmd")

```

See the html generated in the "Results" directory (the location can be changed
and extracted using `results_dir()`)

Let's generate a custom VPC and PPC diagnostic for this run using NMproject's
built in markdown templates.  These just require a simulation control file of
the model to be run.

We can do this using the child functionality. We create a new object m1s.
Start with `m1` and pipe into `child()` function specifying a new run identifier
(here we use the same "m1s" identifier)

```{r sim}

m1s <- m1 %>% child("m1s") %>%  
  # update the $THETA/$OMEGA/$SIGMA values using m1
  update_parameters(m1) %>%
  # use the convert_to_simulation() function to remove $EST and replace with
  # $SIM
  convert_to_simulation(subpr = 10) %>%
  # before running, check the nm_diff (highlight this entire segment, addins ->
  # nm_diff) to ensure automatic control file changes modified m1 in the way
  # intended.
  run_nm()

```

the following templates come are already save in "Scripts" for the demo, but
they are available as templates in `new file` -> `R markdown` -> `from templates`

```{r sim_diags}
 
m1s %>% nm_render("Scripts/basic_ppc.Rmd")
m1s %>% nm_render("Scripts/basic_vpc.Rmd")

```

check "Results to view" and it's good to document your thoughts about how well
the model evalation is so others (including your future self) may follow your
thinking

```{r manual_edit}
# example manual_edit - no run_nm() because will not run this, it's just to
# illustrate how code looks after using the manual_edit addin

m1s_temp <- m1 %>% child("m1s_temp") %>%  
  update_parameters(m1) %>%
  apply_manual_edit("patch-tarjinde-2021-04-20-13-50-09")

# try using nm_diff(m1s_temp, m1s) to see the difference between the two
# approaches.  Note that since we didn't run_nm(), runm1s_temp.mod was never
# written to disk. The control file exists only inside the object

```


```{r child_obs}
# m1 will be used in a future script, so we are saving the object in the
# "Results" directory.  This way it can be reused in other scripts with m1 <-
# readRDS("Results/m1.RDS") without having to rerun this script.

m1 %>% saveRDS("Results/m1.RDS")

# create a child run where we investigate a different TRANS subroutine
m2 <- m1 %>% child("m2") %>%
  subroutine(trans = 2) %>% 
  # see nm_diff app to see if you're happy with the changes being made
  run_nm()


# similarly, we can add coviarates and do many more types of manipulation
m2WT <- m2 %>% child("m2WT") %>%
  add_cov(param = "CL", cov = "WT", state = "linear") %>%
  run_nm()

# ?see manual_edits to see how the manual_edit app works - it can track and
# apply manual control files changes you want to make for when wanting to make
# changes that I haven't made an R function for.  Also check out ?dollar and
# ?text for how to view and write control file control files as segment or in
# their entirety.  ?show_ctl is another way.

# uncomment following to do basic gofs for these too:

# c(m2, m2WT) %>% nm_render("Scripts/basic_gof.Rmd")

```

Body weight not signficant.  K parameterisation better than CL in terms of OFV

```{r summary_rr}

summary_long(c(m2, m2WT), parameters = "all")
rr(c(m2, m2WT))

```

```{r boot}

# here we do a bootstrap. This relies of bootstrap splits being already
# calculated - see the s01 script for how to make them

m1_boot <- m1 %>% 
  make_boot_datasets(samples = 5, overwrite = TRUE)

m1_boot$m <- m1_boot$m %>% run_nm()
wait_finish(m1_boot$m)

m1_boot$m %>% nm_list_render("Scripts/basic_boot.Rmd")


```

```{r xv, eval=FALSE}
# can also do bootstrap cross validation as follows

m1_xv <- m1_boot %>% 
  make_xv_datasets() %>%
  mutate(m_xv = m_xv %>%
           update_parameters(m1_boot$m) %>%
           dollar("EST",
                  "$EST METHOD=IMP INTER EONLY= 1 MAPITER=0 ISAMPLE = 2000 NITER = 10 RANMETHOD=3S2 NOABORT PRINT=1 NSIG=3 SIGL=9"))

m1_xv$m_xv <- m1_xv$m_xv %>% run_nm()
wait_finish(m1_xv$m_xv)

m1_xv$m_xv %>% 
  subset(status(.) %in% "finished") %>% 
  ofv() %>% 
  mean() ## xv-ofv - to be used for xv model selection

```


